---
title: "Neural network training (Preliminary task 3)"
output: html_document
---

This follows a similar structure to `nlp-model-development.R`. 

## Setup

In this section I load the raw data, apply the existing preprocessing pipeline, and create a cleaned version of the data for modeling.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(keras)
library(tensorflow)

source("preprocessing.R")

set.seed(133233)
```

```{r preprocess}
load("../data/claims-raw.RData")

claims_clean <- claims_raw %>%
  parse_data()
```

## Train and test split

Splitting into two different datasets: one is for testing and one for training (with a $20/80$ split respectively).

```{r split}
load("data/claims-clean-task3.RData")

partitions <- claims_clean %>%
  initial_split(prop = 0.8)

train_df <- training(partitions)
test_df  <- testing(partitions)

train_text <- train_df %>% pull(text_clean)
train_labels <- train_df %>%
  pull(bclass) %>%
  as.numeric() - 1

test_text <- test_df %>% pull(text_clean)
test_labels <- test_df %>%
  pull(bclass) %>%
  as.numeric() - 1
```

## Text vectorization layer

I use a Keras text vectorization layer that converts raw text into tf-idf document vectors. The layer is fitted on the training text and then reused across all neural network configurations.

```{r text-vectorization}
preprocess_layer <- layer_text_vectorization(
  standardize = NULL,
  split = 'whitespace',
  ngrams = NULL,
  max_tokens = NULL,
  output_mode = 'tf_idf'
)

preprocess_layer %>% adapt(train_text)
```

## Neural network configurations

Now, I made a small set of candidates for architectures. The difference for each of them is the number of hidden units and layers changes. 

```{r}
install_tensorflow()
# define NN architecture
model1 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 25) %>%
  layer_dropout(0.2) %>%
  layer_dense(1) %>%
  layer_activation(activation = 'sigmoid')

summary(model)
```

I use a helper function that creates and compiles a model given a configuration.

```{r build-model}
build_model <- function(units1, units2 = NA_integer_, dropout = 0.3) {
  model <- keras_model_sequential() %>%
    preprocess_layer() %>%
    layer_dropout(rate = dropout) %>%
    layer_dense(units = units1, activation = "relu")
  
  if (!is.na(units2)) {
    model <- model %>%
      layer_dropout(rate = dropout) %>%
      layer_dense(units = units2, activation = "relu")
  }
  
  model <- model %>%
    layer_dropout(rate = dropout) %>%
    layer_dense(units = 1, activation = "sigmoid")
  
  model %>% compile(
    loss = "binary_crossentropy",
    optimizer = "adam",
    metrics = "binary_accuracy"
  )
  
  model
}


# configure for training
model %>% compile(
  loss = 'binary_crossentropy',
  optimizer = 'adam',
  metrics = 'binary_accuracy'
)

# train
history <- model %>%
  fit(train_text, 
      train_labels,
      validation_split = 0.3,
      epochs = 5)

## CHECK TEST SET ACCURACY HERE

# save the entire model as a SavedModel
save_model_tf(model, "results/example-model")
```
```

## Training and model selection

For each configuration, I train a model with early stopping and record both validation accuracy and test accuracy. The validation split is taken from the training portion, and the test accuracy is computed on the held out test set.

```{r train-loop, message=FALSE, warning=FALSE}
callback_list <- list(
  callback_early_stopping(
    monitor = "val_loss",
    patience = 3,
    restore_best_weights = TRUE
  )
)

results <- vector("list", nrow(configs))

for (i in seq_len(nrow(configs))) {
  cfg <- configs[i, ]
  
  model_i <- build_model(
    units1  = cfg$units1,
    units2  = cfg$units2,
    dropout = cfg$dropout
  )
  
  history_i <- model_i %>%
    fit(
      x = train_text,
      y = train_labels,
      validation_split = 0.3,
      epochs = 20,
      callbacks = callback_list,
      verbose = 0
    )
  
  val_acc <- max(history_i$metrics$val_binary_accuracy)
  
  test_pred_probs <- model_i %>%
    predict(test_text) %>%
    as.numeric()
  
  test_pred_class <- ifelse(test_pred_probs > 0.5, 1, 0)
  test_acc <- mean(test_pred_class == test_labels)
  
  results[[i]] <- list(
    name    = cfg$name,
    val_acc = val_acc,
    test_acc = test_acc,
    model   = model_i,
    history = history_i
  )
}

results_df <- purrr::map_dfr(
  results,
  ~ tibble(
    name    = .x$name,
    val_acc = .x$val_acc,
    test_acc = .x$test_acc
  )
)

results_df
```

