---
title: "Neural network training (Preliminary task 3)"
output: html_document
---

This follows a similar structure to `nlp-model-development.R`. 

## Setup

In this section I load the raw data, apply the existing preprocessing pipeline, and create a cleaned version of the data for modeling.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(keras)
library(tensorflow)

source("preprocessing.R")

set.seed(133233)
```

```{r preprocess}
load("../data/claims-raw.RData")

claims_clean <- claims_raw %>%
  parse_data()
```

## Train and test split

Splitting into two different datasets: one is for testing and one for training (with a $20/80$ split respectively).

```{r split}
load("data/claims-clean-task3.RData")

partitions <- claims_clean %>%
  initial_split(prop = 0.8)

train_df <- training(partitions)
test_df  <- testing(partitions)

train_text <- train_df %>% pull(text_clean)
train_labels <- train_df %>%
  pull(bclass) %>%
  as.numeric() - 1

test_text <- test_df %>% pull(text_clean)
test_labels <- test_df %>%
  pull(bclass) %>%
  as.numeric() - 1
```

## Text vectorization layer

I use a Keras text vectorization layer that converts raw text into tf-idf document vectors. The layer is fitted on the training text and then reused across all neural network configurations.

```{r text-vectorization}
preprocess_layer <- layer_text_vectorization(
  standardize = NULL,
  split = 'whitespace',
  ngrams = NULL,
  max_tokens = NULL,
  output_mode = 'tf_idf'
)

preprocess_layer %>% adapt(train_text)
```

## Neural network configurations

Now, I made a small set of candidates for architectures. The difference for each of them is the number of hidden units and layers changes. 

```{r}
# model 1: shallow model with small hidden layer (baseline)
model1 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 25, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(1, activation = "sigmoid")

summary(model1)



# model 2: wider single layer
model2 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(1, activation = "sigmoid")

summary(model2)



# model 3: two hidden layers to test deeper learning
model3 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(1, activation = "sigmoid")

summary(model3)



# model 4: smaller model with stronger dropout
model4 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.5) %>% 
  layer_dense(units = 32, activation = "relu") %>%
  layer_dropout(0.5) %>%
  layer_dense(1, activation = "sigmoid")

summary(model4)
```

## Training and model selection

Now, for each model, we will configure the model and train it:
```{r}
# configure
model1 %>% compile(
  loss = "binary_crossentropy",
  optimizer = "adam",
  metrics = "binary_accuracy"
)

# train
history1 <- model1 %>%
  fit(
    train_text,
    train_labels,
    validation_split = 0.3,
    epochs = 5
  )


# configure
model2 %>% compile(
  loss = "binary_crossentropy",
  optimizer = "adam",
  metrics = "binary_accuracy"
)

# train
history2 <- model2 %>%
  fit(
    train_text,
    train_labels,
    validation_split = 0.3,
    epochs = 5
  )


# configure
model3 %>% compile(
  loss = "binary_crossentropy",
  optimizer = "adam",
  metrics = "binary_accuracy"
)

# train
history3 <- model3 %>%
  fit(
    train_text,
    train_labels,
    validation_split = 0.3,
    epochs = 5
  )


# configure
model4 %>% compile(
  loss = "binary_crossentropy",
  optimizer = "adam",
  metrics = "binary_accuracy"
)

# train
history4 <- model4 %>%
  fit(
    train_text,
    train_labels,
    validation_split = 0.3,
    epochs = 5
  )
```


## Testing accuracy for each model

Now, we test the accuracy of each model trained. 
```{r}
# model 1
pred1 <- model1 %>%
  predict(test_text) %>%
  as.numeric()
class1 <- ifelse(pred1 > 0.5, 1, 0)
acc1 <- mean(class1 == test_labels)


# model 2
pred2 <- model2 %>%
  predict(test_text) %>%
  as.numeric()
class2 <- ifelse(pred2 > 0.5, 1, 0)
acc2 <- mean(class2 == test_labels)


# model 3
pred3 <- model3 %>%
  predict(test_text) %>%
  as.numeric()
class3 <- ifelse(pred3 > 0.5, 1, 0)
acc3 <- mean(class3 == test_labels)


# model 4
pred4 <- model4 %>%
  predict(test_text) %>%
  as.numeric()
class4 <- ifelse(pred4 > 0.5, 1, 0)
acc4 <- mean(class4 == test_labels)


# summary table
tibble(
  model = c("model1", "model2", "model3", "model4"),
  test_accuracy = c(acc1, acc2, acc3, acc4)
)
```


## Saving the best model

```{r}
save_model_tf(model1, "../results/model1")
save_model_tf(model2, "../results/model2")
save_model_tf(model3, "../results/model3")
save_model_tf(model4, "../results/model4")
```

