---
title: "Neural network training (Preliminary task 3)"
output: html_document
---

This follows a similar structure to `nlp-model-development.R`. 

## Setup

In this section I load the raw data, apply the existing preprocessing pipeline, and create a cleaned version of the data for modeling.

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(keras)
library(tensorflow)

source("preprocessing.R")

set.seed(133233)
```

```{r preprocess}
load("../data/claims-raw.RData")

claims_clean <- claims_raw %>%
  parse_data()
```

## Train and test split

Splitting into two different datasets: one is for testing and one for training (with a $20/80$ split respectively).

```{r split}
load("data/claims-clean-task3.RData")

partitions <- claims_clean %>%
  initial_split(prop = 0.8)

train_df <- training(partitions)
test_df  <- testing(partitions)

train_text <- train_df %>% pull(text_clean)
train_labels <- train_df %>%
  pull(bclass) %>%
  as.numeric() - 1

test_text <- test_df %>% pull(text_clean)
test_labels <- test_df %>%
  pull(bclass) %>%
  as.numeric() - 1
```

## Text vectorization layer

I use a Keras text vectorization layer that converts raw text into tf-idf document vectors. The layer is fitted on the training text and then reused across all neural network configurations.

```{r text-vectorization}
preprocess_layer <- layer_text_vectorization(
  standardize = NULL,
  split = 'whitespace',
  ngrams = NULL,
  max_tokens = NULL,
  output_mode = 'tf_idf'
)

preprocess_layer %>% adapt(train_text)
```

## Neural network configurations

Now, I made a small set of candidates for architectures. The difference for each of them is the number of hidden units and layers changes. 

```{r}
# model 1: shallow model with small hidden layer (baseline)
model1 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.2) %>%
  layer_dense(units = 25, activation = "relu") %>%
  layer_dropout(0.2) %>%
  layer_dense(1, activation = "sigmoid")

summary(model1)



# model 2: wider single layer
model2 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(1, activation = "sigmoid")

summary(model2)



# model 3: two hidden layers to test deeper learning
model3 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 64, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dropout(0.3) %>%
  layer_dense(1, activation = "sigmoid")

summary(model3)



# model 4: smaller model with stronger dropout
model4 <- keras_model_sequential() %>%
  preprocess_layer() %>%
  layer_dropout(0.5) %>% 
  layer_dense(units = 32, activation = "relu") %>%
  layer_dropout(0.5) %>%
  layer_dense(1, activation = "sigmoid")

summary(model4)
```


